LCS Tool: A Computational Platform for Lagrangian Coherent Structures

### 2014-12-11 main.tex
Remove unpublished reference
Remove & revise FIXMEs

Next we consider positions of material lines at time $t_0$ that prevail as most repelling or attracting material lines (or \emph{hyperbolic LCSs}) over a time interval $[t_0,t] \subset [t_-,t_+]$.
\textcite{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
\begin{gather*}
Q(\gamma) = \frac{1}{\sigma} \int_0^\sigma \frac{\langle r'(s),D_{t_0}^t(r(s)) r'(s)\rangle}{\sqrt{\langle r'(s),C_{t_0}^t(r(s)) r'(s)\rangle\langle r'(s),r'(s)\rangle}}\dif s,\\
D_{t_0}^t = \frac12 [C_{t_0}^t \Omega - \Omega C_{t_0}^t],
\end{gather*}
% FIXME Clarify what is meant by _closed_ material lines
obtained by averaging the Lagrangian shear arising over $[t_0,t]$ along closed material lines parametrized as $r(s)$ with $s \in [0,\sigma]$.
More precisely, hyperbolic LCSs are stationary curves of $Q(\gamma)$ with respect to fixed-endpoint perturbations.
We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \textcite{farazmand14:_shearless}).

Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field.
Repelling LCSs (\emph{shrink lines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_1(r),
\label{eq:shrink line}
\end{equation}
and attracting LCSs (\emph{stretch lines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_2(r).
\label{eq:stretch line}
\end{equation}

\begin{sloppypar}
Shrink lines and stretch lines coincide with the null geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$. For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting orbits of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
\end{sloppypar}

@@

\begin{table}
\begin{enumerate}
\item Define a velocity field, \lstinline!dx = derivative(t,x,p)!
\item Compute Cauchy-Green strain tensor invariants, \lstinline![v,d] = eig_cgStrain(derivative)!
\item Compute $\lambda$-lines, \lstinline![etaPos,etaNeg] = lambda_line(v,d,lambda)!
\item Define Poincare sections, \lstinline!ps!
% FIXME function call below does not reflect LCS Tool usage (e.g. poincare_close_orbit_multi)
\item Detect elliptic LCSs, \lstinline!ellipticLcs = poincare_closed_orbit(etaPos,etaNeg,ps)!
\item Define a local maximization/minimization distance for hyperbolic LCSs \lstinline!lmd!
% FIXME function call below does not reflect LCS Tool usage (e.g. remove_strain_in_shear)
% FIXME seed_curves_from_lambda_max is a poor name for a function that detects hyperbolic LCSs
\item Detect hyperbolic LCSs, \lstinline!hyperbolicLcs = seed_curves_from_lambda_max(v,d,ellipticLcs,lmd)!
\end{enumerate}
\caption{Overview of sequence of computations to detect LCSs with LCS Tool functions.}
\label{t:LCS Tool overview}
\end{table}

@@

% FIXME Need to update table to account for lambda range
\begin{table*}
\begin{tabular}{|c|p{.7\textwidth}|}
\hline
\multicolumn{2}{|p{.95\textwidth}|}
{\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\tabularnewline
\hline \hline
\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
\hline
\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\tabularnewline
\hline
\lstinline!lambda! & scalar lambda value in \cref{eq:eta}\tabularnewline
\hline \hline
\multicolumn{2}{|p{.95\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\tabularnewline
\hline \hline
\lstinline!domain! & array to define flow domain\tabularnewline
\hline
\lstinline!resolution! & $1 \times 2$ array to define main grid resolution for Cauchy-Green strain tensor\tabularnewline
\hline
\lstinline!shearline! & structure of arrays of $\eta_+$ and $\eta_-$ values on main grid\tabularnewline
\hline
\lstinline!PSList! & user-defined structure for Poincare section end-points, number of $\lambda$-lines launched from Poincare section, and maximum closed $\lambda$-line length\tabularnewline
\hline
\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default: 5.\tabularnewline
\hline
\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare map. Default: $10^{-2}$.\tabularnewline
\hline
\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify $\lambda$-line integration parameters\tabularnewline
\hline
\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default: \lstinline![false,false]!.\tabularnewline
\hline
\end{tabular}
\caption{Syntax for LCS Tool elliptic LCS functions.}
\label{t:Elliptic LCS functions}
\end{table*}

@@

% FIXME Clarify use of the term "orbit"
As discussed in \cref{sec:Hyperbolic LCSs}, hyperbolic LCS positions at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:shrink line} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretch line} (attracting LCSs).
By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
We identify the strongest repelling shrink lines as those crossing a local maximum of the $\lambda_2(x_0)$ field.
Similarly, we identify the strongest attracting stretch lines as those crossing a local minimum of the $\lambda_1(x_0)$ field.
These local maxima and minima of the appropriate $\lambda_i(x_0)$ eigenvalue field can be thought of as the extensions of the concept of saddle points to the present finite-time, temporally aperiodic flow setting.

>>>>

Next we consider positions of material lines at time $t_0$ that prevail as most repelling or attracting material lines (or \emph{hyperbolic LCSs}) over a time interval $[t_0,t] \subset [t_-,t_+]$.
\textcite{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
\begin{gather*}
Q(\gamma) = \frac{1}{\sigma} \int_0^\sigma \frac{\langle r'(s),D_{t_0}^t(r(s)) r'(s)\rangle}{\sqrt{\langle r'(s),C_{t_0}^t(r(s)) r'(s)\rangle\langle r'(s),r'(s)\rangle}}\dif s,\\
D_{t_0}^t = \frac12 [C_{t_0}^t \Omega - \Omega C_{t_0}^t],
\end{gather*}
obtained by averaging the Lagrangian shear arising over $[t_0,t]$ along material lines parametrized as $r(s)$ with $s \in [0,\sigma]$.
More precisely, hyperbolic LCSs are stationary curves of $Q(\gamma)$ with respect to fixed-endpoint perturbations.
We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \textcite{farazmand14:_shearless}).

Solutions to this variational problem turn out to be curves of the $\xi_1$ or $\xi_2$ eigenvector field.
Repelling LCSs (\emph{shrink lines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_1(r),
\label{eq:shrink line}
\end{equation}
and attracting LCSs (\emph{stretch lines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_2(r).
\label{eq:stretch line}
\end{equation}

\begin{sloppypar}
Shrink lines and stretch lines coincide with the null geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting curves of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
\end{sloppypar}

@@

\begin{table}
\begin{enumerate}
\item Define a velocity field, \lstinline!dx = derivative(t,x,p)!
\item Compute Cauchy-Green strain tensor invariants, \lstinline![v,d] = eig_cgStrain(derivative)!
% FIXME lambda_line and poincare_closed_orbit have been removed from demo scripts. Now use poincare_closed_orbit_range and elliptic_lcs.
\item Compute $\lambda$-lines, \lstinline![etaPos,etaNeg] = lambda_line(v,d,lambda)!
\item Define Poincare sections, \lstinline!ps!
\item Detect elliptic LCSs, \lstinline!ellipticLcs = poincare_closed_orbit(etaPos,etaNeg,ps)!
\item Define a local maximization/minimization distance for hyperbolic LCSs \lstinline!lmd!
% FIXME function call below does not reflect LCS Tool usage (e.g. remove_strain_in_shear)
% FIXME seed_curves_from_lambda_max is a poor name for a function that detects hyperbolic LCSs
\item Detect hyperbolic LCSs, \lstinline!hyperbolicLcs = seed_curves_from_lambda_max(v,d,ellipticLcs,lmd)!
\end{enumerate}
\caption{Overview of sequence of computations to detect LCSs with LCS Tool functions.}
\label{t:LCS Tool overview}
\end{table}

@@

% FIXME lambda_line and poincare_closed_orbit have been removed from demo scripts. Now use poincare_closed_orbit_range and elliptic_lcs.
\begin{table*}
\begin{tabular}{|c|p{.7\textwidth}|}
\hline
\multicolumn{2}{|p{.95\textwidth}|}
{\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\tabularnewline
\hline \hline
\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
\hline
\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\tabularnewline
\hline
\lstinline!lambda! & scalar lambda value in \cref{eq:eta}\tabularnewline
\hline \hline
\multicolumn{2}{|p{.95\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\tabularnewline
\hline \hline
\lstinline!domain! & array to define flow domain\tabularnewline
\hline
\lstinline!resolution! & $1 \times 2$ array to define main grid resolution for Cauchy-Green strain tensor\tabularnewline
\hline
\lstinline!shearline! & structure of arrays of $\eta_+$ and $\eta_-$ values on main grid\tabularnewline
\hline
\lstinline!PSList! & user-defined structure for Poincare section end-points, number of $\lambda$-lines launched from Poincare section, and maximum closed $\lambda$-line length\tabularnewline
\hline
\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default: 5.\tabularnewline
\hline
\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare map. Default: $10^{-2}$.\tabularnewline
\hline
\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify $\lambda$-line integration parameters\tabularnewline
\hline
\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default: \lstinline![false,false]!.\tabularnewline
\hline
\end{tabular}
\caption{Syntax for LCS Tool elliptic LCS functions.}
\label{t:Elliptic LCS functions}
\end{table*}

@@

As discussed in \cref{sec:Hyperbolic LCSs}, hyperbolic LCS positions at time $t_0$ are found as the strongest repelling curves of the vector field \cref{eq:shrink line} (repelling LCSs), and strongest attracting curves of the vector field \cref{eq:stretch line} (attracting LCSs).
By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
We identify the strongest repelling shrink lines as those crossing a local maximum of the $\lambda_2(x_0)$ field.
Similarly, we identify the strongest attracting stretch lines as those crossing a local minimum of the $\lambda_1(x_0)$ field.
These local maxima and minima of the appropriate $\lambda_i(x_0)$ eigenvalue field can be thought of as the extensions of the concept of saddle points to the present finite-time, temporally aperiodic flow setting.
